---
title: "SQL"
output: html_document
---
## Filtering Columns and Rows

With SQL, you can filter columns and rows by using SELECT for columns and WHERE from rows.  Let's look at an example using the Lahman database. So we should first load that database, along with the package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```
Suppose now I would like to see the homerun totals for the 1927 Yankees.  I could write the following query
query:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```


Now we will find any instance where a player had more than 40 homeruns but less than 60 strikeouts.

```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting 
WHERE HR>40 and SO<60"
sqldf(query)
```

Let's look at instances of Phillies in the 1970s hitting more than 30 homeruns.

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID<=1979 and yearID>=1970"
sqldf(query)
```



## ORDER BY

With SQL, you can order the rows of your output. For instance, suppose you want to see every instance of a player hitting more than 50 homeruns, but you would like the players with the most homeruns to be at the top.  We would do the following:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

